---
title: "xu41_42"
output: html_document
---


Start R (by typing ‘R’ at the command line) and load some R libraries as follows:

## step1 
```{r}

library("rlang")
library("Seurat");
library("sctransform");
library("dplyr");
library("RColorBrewer");
library("ggthemes");
library("ggplot2");
library("cowplot");
library("data.table");
library("xfun");
library("WGCNA")

```


```{r}
## Step 2: find the origial CR path for three whole ear samples

setwd("C:/Users/Xiaosa Xu/Desktop/Seurat_2020/xu41_ap1cal")

setwd("C:/Users/Xiaosa Xu/Desktop/Seurat_2020/xu42_ap1cal")



## Step 3: Read in the feature-barcode matrices generated by the cellranger pipeline.

memory.limit(size=50000000)

samples=c("xu41","xu42")

data.10x = list()

data.10x[[1]] <- Read10X(data.dir = "C:/Users/Xiaosa Xu/Desktop/Seurat_2020/xu41_ap1cal/raw_feature_bc_matrix");

data.10x[[2]] <- Read10X(data.dir = "C:/Users/Xiaosa Xu/Desktop/Seurat_2020/xu42_ap1cal/raw_feature_bc_matrix");



## Step 4: Convert each feature-barcode matrix to a Seurat object.

ap1cal.list = list(); # First create an empty list to hold the Seurat objects

ap1cal.list[[1]] = CreateSeuratObject(counts = data.10x[[1]], min.cells=3, min.features=200, project=samples[1]);

ap1cal.list[[1]][["DataSet"]] = samples[1];

ap1cal.list[[2]] = CreateSeuratObject(counts = data.10x[[2]], min.cells=3, min.features=200, project=samples[2]);

ap1cal.list[[2]][["DataSet"]] = samples[2];



## Step 5: Optional, remove the raw data to save space:

rm(data.10x);


### subset each library before merge or integrate


ap1cal.list[[1]] <- subset(ap1cal.list[[1]], subset = nCount_RNA > 1000 & nFeature_RNA > 500 & nCount_RNA < 100000 & nFeature_RNA < 12000)

ap1cal.list[[2]] <- subset(ap1cal.list[[2]], subset = nCount_RNA > 1000 & nFeature_RNA > 500 & nCount_RNA < 100000 & nFeature_RNA < 12000)


```

```{r}
## Step 8-9:for seurat intergration, firstly need to log normalize and then find variables. 
# Filter each sample separately and normalize (using same method for all samples)

## Step 8:logNormalize

ap1cal.list[[1]]<- NormalizeData(ap1cal.list[[1]], normalization.method = "LogNormalize", scale.factor = 10000)

ap1cal.list[[2]]<- NormalizeData(ap1cal.list[[2]], normalization.method = "LogNormalize", scale.factor = 10000)


##Step 9:Find variables


ap1cal.list[[1]]<- FindVariableFeatures(ap1cal.list[[1]], selection.method = "vst", nfeatures = 2000)

ap1cal.list[[2]]<- FindVariableFeatures(ap1cal.list[[2]], selection.method = "vst", nfeatures = 2000)


```



```{r}
features <- SelectIntegrationFeatures(object.list = ap1cal.list)

```



## Step 10:Next Integration of 3 replicates 

# find anchors

```{r}
anchors <- FindIntegrationAnchors(object.list = ap1cal.list, anchor.features = features)
```


# Step 11:Integrate data


```{r}

ap1cal.int <- IntegrateData(anchorset = anchors)

```

```{r}
DefaultAssay(ap1cal.int)<-"integrated"


# Step 13:Run the standard workflow for visualization and clustering
ap1cal.int <- ScaleData(ap1cal.int)
ap1cal.int <- RunPCA(ap1cal.int)


```


```{r}
## Determine the ‘dimensionality’ of the dataset

ElbowPlot(ap1cal.int, ndims = 40)

## Elbow plot: quantitative approach
## https://hbctraining.github.io/scRNA-seq/lessons/elbow_plot_metric.html

## Based on ElbowPlot(xu19, ndims = 40) plot, we could roughly determine the majority of the variation by where the elbow occurs (touches the ground) to be between PC12-PC16. While this gives us a good rough idea of the number of PCs needed to be included, a more quantitative approach may be a bit more reliable. We can calculate where the principal components start to elbow by taking the larger value of:

## 1. The point where the principal components only contribute 5% of standard deviation and the principal components cumulatively contribute 90% of the standard deviation.
## 2. The point where the percent change in variation between the consecutive PCs is less than 0.1%.

## We will start by calculating the first metric:
# Determine percent of variation associated with each PC
pct <- ap1cal.int[["pca"]]@stdev / sum(ap1cal.int[["pca"]]@stdev) * 100

# Calculate cumulative percents for each PC
cumu <- cumsum(pct)

# Determine which PC exhibits cumulative percent greater than 90% and % variation associated with the PC as less than 5
co1 <- which(cumu > 90 & pct < 5)[1]

co1

## [1] 43

## The first metric returns PC43 as the PC matching these requirements. Let’s check the second metric, which identifies the PC where the percent change in variation between consecutive PCs is less than 0.1%:

# Determine the difference between variation of PC and subsequent PC
co2 <- sort(which((pct[1:length(pct) - 1] - pct[2:length(pct)]) > 0.1), decreasing = T)[1] + 1

# last point where change of % of variation is more than 0.1%.
co2

## [1] 15

## This second metric returns PC14. Usually, we would choose the minimum of these two metrics as the PCs covering the majority of the variation in the data.

# Minimum of the two calculation
pcs <- min(co1, co2)

pcs


```



```{r}
ap1cal.int <- RunUMAP(ap1cal.int, reduction = "pca", dims = 1:21)

ap1cal.int <- FindNeighbors(ap1cal.int, reduction = "pca", dims = 1:21)

ap1cal.int <- FindClusters(ap1cal.int, resolution = 0.78)

```

```{r}
DimPlot(ap1cal.int, reduction = "umap", pt.size = 1, label = TRUE)

```


```{r}
# plot number of gene per cell on this plot. 
# https://github.com/satijalab/seurat/issues/2192
VlnPlot(ap1cal.int, features = c("nFeature_RNA"), pt.size = 0.5)
```



```{r}
# plot number of transcript/UMI per cell on this plot. 
# https://github.com/satijalab/seurat/issues/2192
VlnPlot(ap1cal.int, features = c("nCount_RNA"), pt.size = 0.5)
```



```{r}

VlnPlot(ap1cal.int, features = c("nFeature_RNA", "nCount_RNA"), ncol=2)

```



```{r}
ID	Name	Domain
AT2G42840	PDF1	Epidermis
AT4G21750	ATML1	Epidermis
AT2G27250	CLV3	Stem cell
AT2G17950	WUS	OC
AT1G62360	STM	Meristem
AT1G76420   CUC
AT5G61850	LEAFY	FM
AT5G03840	TFL	RM?
AT4G00180	YAB3	FIL
AT5G10400	H3.1	S phase 
AT2G28740	HIS4	S phase 
AT1G76540	CDKB2;1	G2/M phase
AT1G20930	CDKB2;2	G2/M phase
AT1G44110	CYCA1;1	G2/M phase
AT4G37490	CYCB1;1 	G2/M phase
AT3G53980	Protease inhibitor	Vas
AT5G50790	Nodulin MtN3 family protein	Vas
AT2G02130	LCR68	Vas
AT5G61480	PXY	Xylem
AT3G25710	TMO5	Xylem
AT1G68810	T5L1/ABS5	Xylem
AT1G20850	XCP2	Trachery Element
AT4G35350	XCP1	Trachery Element
AT1G79430	APL	Phloem
AT3G28455	CLE25	Phloem
AT4G19840	AN5/ATPP2-A1	Companion Cell
AT1G35750	PUM10	CLV3
AT5G65510	AIL7	CLV3
AT5G06270	Cyclin-like	CLV3-L1
AT5G06300	Carboxy-lyase	CLV3
AT5G07930	MCT2	CLV3
AT1G04880	HMG	CLV3-L1
AT4G17710	HDG4	CLV3
AT1G37140	MCT1	CLV3
AT1G70560	Alliinase	CLV3-and-vas
AT3G59270	Syntaxin	CLV3-and-WUS
AT54G39410	WRKY13	CLV3-and-WUS
AT5G28650	WRKY 74	CLV3-and-WUS-and Vas
AT5G56970	CKX3	CLV3-and-WUS
AT1G21460	Nodulin MtN3 family protein	CLV3-and-WUS
AT1G26680	B3 TF	WUS-and-CLV3
AT1G76110	HMG	WUS-and-CLV3-and-whoel Central Zone
AT4G34590	ATBZIP11	WUS
AT4G09760	Choline Kinase	WUS only very pure
AT5G54510	DFL1	WUS-and-Vas
AT4G39480	CYT p450	WUS
AT4G21690	GA3OX3	FIL
AT5G65410	ATHB25	FIL
AT4G00180	YAB3	FIL
AT2G45190	FIL/YAB1	FIL
AT5G16560	KAN1	Abaxial
AT1G32240	KAN2	Abaxial
AT4G17695	KAN3	Abaxial
AT5G42630	KAN4	Abaxial
AT5G60690	REV	Adaxial
AT2G34710	PHB	Adaxial
AT1G30490	PHV	Adaxial
AT2G40360	BOP1	Adaxial


DefaultAssay(ap1cal.int)<-"RNA"

FeaturePlot(ap1cal.int, features = c("AT2G42840","AT4G21750","AT2G27250","AT2G17950","AT1G62360","AT5G61850","AT5G03840","AT4G00180","AT5G10400","AT2G28740","AT1G76540","AT1G20930","AT1G44110","AT4G37490","AT3G53980","AT5G50790","AT2G02130","AT5G61480","AT3G25710","AT1G68810","AT1G20850","AT4G35350","AT1G79430","AT3G28455","AT4G19840","AT1G35750","AT5G65510","AT5G06270","AT5G06300","AT5G07930","AT1G04880","AT4G17710","AT1G37140","AT1G70560","AT3G59270","AT54G39410","AT5G28650","AT5G56970","AT1G21460","AT1G26680","AT1G76110","AT4G34590","AT4G09760","AT5G54510","AT4G39480","AT4G21690","AT5G65410","AT4G00180","AT2G45190","AT5G16560","AT1G32240","AT4G17695","AT5G42630","AT5G60690","AT2G34710","AT1G30490","AT2G40360"))



VlnPlot(ap1cal.int, features = c("AT2G42840","AT4G21750","AT2G27250","AT2G17950","AT1G62360","AT5G61850","AT5G03840","AT4G00180","AT5G10400","AT2G28740","AT1G76540","AT1G20930","AT1G44110","AT4G37490","AT3G53980","AT5G50790","AT2G02130","AT5G61480","AT3G25710","AT1G68810","AT1G20850","AT4G35350","AT1G79430","AT3G28455","AT4G19840","AT1G35750","AT5G65510","AT5G06270","AT5G06300","AT5G07930","AT1G04880","AT4G17710","AT1G37140","AT1G70560","AT3G59270","AT54G39410","AT5G28650","AT5G56970","AT1G21460","AT1G26680","AT1G76110","AT4G34590","AT4G09760","AT5G54510","AT4G39480","AT4G21690","AT5G65410","AT4G00180","AT2G45190","AT5G16560","AT1G32240","AT4G17695","AT5G42630","AT5G60690","AT2G34710","AT1G30490","AT2G40360"))



```





```{r}
# Step 18: with cluster# and cell#
cell.num <- table(ap1cal.int@active.ident)
ClusterLabels = paste("Cluster", names(cell.num), paste0("(n = ", cell.num, ")"))
ClusterBreaks = names(cell.num)

new.cluster.ids <- c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11","12","13","14","15")
names(new.cluster.ids) <- levels(ap1cal.int)
scrna.int<-RenameIdents(ap1cal.int, new.cluster.ids)
DimPlot(ap1cal.int,reduction = "umap", label = TRUE, pt.size = 1, label.size = 10) + scale_color_discrete(breaks= ClusterBreaks, labels=ClusterLabels)


```




## Step 19:plot UMIs and Genes on the UMAP

feature.pal = rev(colorRampPalette(brewer.pal(11,"Spectral"))(50))

FeaturePlot(object = ap1cal.int, features = c("nCount_RNA"), cols = feature.pal, pt.size=1, reduction = "umap") + theme(axis.title.x=element_blank(),axis.title.y=element_blank(),axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank())

FeaturePlot(object = ap1cal.int, features = c("nFeature_RNA"), cols = feature.pal, pt.size=1, reduction = "umap") + theme(axis.title.x=element_blank(),axis.title.y=element_blank(),axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank())



```{r}
## Step 20:save 

saveRDS(ap1cal.int, file = "ap1cal.int_UMI-1000-100000_Gene-1000-12000.rds")

### in future, when open this datase

ap1cal.int <- readRDS(file = "ap1cal.int_UMI-1000-100000_Gene-1000-12000.rds")


## Step 20:save 

saveRDS(ap1cal.int, file = "ap1cal.int_UMI-1000-100000_Gene-1000-12000_res0.78.rds")

### in future, when open this datase

ap1cal.int <- readRDS(file = "ap1cal.int_UMI-1000-100000_Gene-1000-12000_res0.78.rds")


```





```{r}
### split the UMAPs by samples 

DimPlot(ap1cal.int, reduction = "umap", split.by = "orig.ident", pt.size = 1)


### group the UMAPs by samples. 

DimPlot(ap1cal.int, reduction = "umap", group.by = "orig.ident", pt.size = 1)

```



```{r}
# Step 24 for exporting table of coexpression value(pearson correlation value) for gene of interest, 

matrix <- ap1cal.int$RNA@data
matrix_mod<-as.matrix(matrix)

## for any gene of interest. Pearson correlation. 
gene<-as.numeric(matrix_mod["AT2G27250",])
correlations<-apply(matrix_mod,1,function(x){cor(gene,x)})
write.table(correlations, file = "ap1cal.int.Dec.CLV3_Pearson_correlation.txt")


## for any gene of interest. Spearman correlation. 
gene<-as.numeric(matrix_mod["AT2G27250",])
correlations<-apply(matrix_mod,1,function(x){cor(gene,x, method = "spearman")})
write.table(correlations, file = "ap1cal.int.Dec.CLV3_Spearman_correlation.txt")


# for generating the distribution and plot any gene's position on the distribution

ap1cal=read.table("ap1cal.int.Dec.CLV3_Pearson_correlation.txt")

f=ap1cal[,1]!=1

hist(ap1cal[f,1],breaks = 1000,main = "ap1cal_CLV3_Spearman_correlation",xlab="Coexpression")

# to plot position of WUS1 on CLE7 coexpression's distribution

abline(v=ap1cal["gene ID",1],col="red",lwd=2,lty=2)


```


```{r}
## Generate cluster markers

DefaultAssay(ap1cal.int)<-"integrated"

object5.markers <- FindAllMarkers(object = ap1cal.int, features = VariableFeatures(ap1cal.int, assay = 'integrated'),
                                  only.pos = TRUE, 
                                  min.pct = 0.25, 
                                  logfc.threshold = 0.25, 
                                  assay = 'integrated', 
                                  slot = c('data'))

write.table(object5.markers, file=sprintf("ap1cal.int_minpct0.25.Wilcox.xls"), quote=FALSE, sep="\t", row.names=FALSE);
```
