---
title: "xu16_19_29_32"
output: html_document
---


Start R (by typing ‘R’ at the command line) and load some R libraries as follows:

Introduction to scRNA-seq integration

Compiled: January 11, 2022
Source: vignettes/integration_introduction.Rmd

Introduction to scRNA-seq integration

The joint analysis of two or more single-cell datasets poses unique challenges. In particular, identifying cell populations that are present across multiple datasets can be problematic under standard workflows. Seurat v4 includes a set of methods to match (or ‘align’) shared cell populations across datasets. These methods first identify cross-dataset pairs of cells that are in a matched biological state (‘anchors’), can be used both to correct for technical differences between datasets (i.e. batch effect correction), and to perform comparative scRNA-seq analysis of across experimental conditions.

Below, we demonstrate methods for scRNA-seq integration as described in Stuart*, Butler* et al, 2019 to perform a comparative analysis of human immune cells (PBMC) in either a resting or interferon-stimulated state.

Integration goals
The following tutorial is designed to give you an overview of the kinds of comparative analyses on complex cell types that are possible using the Seurat integration procedure. Here, we address a few key goals:

Create an ‘integrated’ data assay for downstream analysis
Identify cell types that are present in both datasets
Obtain cell type markers that are conserved in both control and stimulated cells
Compare the datasets to find cell-type specific responses to stimulation


## https://satijalab.org/seurat/articles/integration_introduction.html


## Load all packages
```{r}

library("rlang")
library("Seurat");
library("sctransform");
library("dplyr");
library("RColorBrewer");
library("ggthemes");
library("ggplot2");
library("cowplot");
library("data.table");
library("scales")


## for slingshot 

library(tidyverse)
library(slingshot)
library(Seurat)
library(tradeSeq)
library(S4Vectors)
library(SingleCellExperiment)


library(slingshot)
library(BUSpaRse)
library(tidyverse)
library(tidymodels)
library(Seurat)
library(scales)
library(viridis)
library(Matrix)


```


## Set up Seurat object
```{r}
## Find the origial CR path

setwd("C:/Users/Xiaosa Xu/Desktop/Seurat_2020/xu16_2017_cus_CR310_S150_L151")

setwd("C:/Users/Xiaosa Xu/Desktop/Seurat_2020/xu19_2017_cus_CR310_S150_L151")

setwd("C:/Users/Xiaosa Xu/Desktop/Seurat_2020/xu29_B73_tip")

setwd("C:/Users/Xiaosa Xu/Desktop/Seurat_2020/xu32_B73_tip")


## Step 3: Read in the feature-barcode matrices generated by the cellranger pipeline.

memory.limit(size=50000000)

samples=c("xu16","xu19","xu29","xu32")

data.10x = list()

data.10x[[1]] <- Read10X(data.dir = "C:/Users/Xiaosa Xu/Desktop/Seurat_2020/xu16_2017_cus_CR310_S150_L151/raw_feature_bc_matrix");

data.10x[[2]] <- Read10X(data.dir = "C:/Users/Xiaosa Xu/Desktop/Seurat_2020/xu19_2017_cus_CR310_S150_L151/raw_feature_bc_matrix");

data.10x[[3]] <- Read10X(data.dir = "C:/Users/Xiaosa Xu/Desktop/Seurat_2020/xu29_B73_tip/raw_feature_bc_matrix");

data.10x[[4]] <- Read10X(data.dir = "C:/Users/Xiaosa Xu/Desktop/Seurat_2020/xu32_B73_tip/raw_feature_bc_matrix");


## Convert each feature-barcode matrix to a Seurat object.

B73_ear_tip.list = list(); # First create an empty list to hold the Seurat objects

B73_ear_tip.list[[1]] = CreateSeuratObject(counts = data.10x[[1]], min.cells=3, min.features=200, project=samples[1]);

B73_ear_tip.list[[1]][["DataSet"]] = samples[1];

B73_ear_tip.list[[2]] = CreateSeuratObject(counts = data.10x[[2]], min.cells=3, min.features=200, project=samples[2]);

B73_ear_tip.list[[2]][["DataSet"]] = samples[2];

B73_ear_tip.list[[3]] = CreateSeuratObject(counts = data.10x[[3]], min.cells=3, min.features=200, project=samples[3]);

B73_ear_tip.list[[3]][["DataSet"]] = samples[3];

B73_ear_tip.list[[4]] = CreateSeuratObject(counts = data.10x[[4]], min.cells=3, min.features=200, project=samples[4]);

B73_ear_tip.list[[4]][["DataSet"]] = samples[4];


## Optional, remove the raw data to save space:

rm(data.10x);


### Subset each library before merge or integrate


B73_ear_tip.list[[1]] <- subset(B73_ear_tip.list[[1]], subset = nCount_RNA > 5000 & nFeature_RNA > 1000 & nCount_RNA < 150000 & nFeature_RNA < 15000)

B73_ear_tip.list[[2]] <- subset(B73_ear_tip.list[[2]], subset = nCount_RNA > 5000 & nFeature_RNA > 1000 & nCount_RNA < 150000 & nFeature_RNA < 15000)

B73_ear_tip.list[[3]] <- subset(B73_ear_tip.list[[3]], subset = nCount_RNA > 5000 & nFeature_RNA > 1000 & nCount_RNA < 150000 & nFeature_RNA < 15000)

B73_ear_tip.list[[4]] <- subset(B73_ear_tip.list[[4]], subset = nCount_RNA > 5000 & nFeature_RNA > 1000 & nCount_RNA < 150000 & nFeature_RNA < 15000)


## Before seurat intergration, filter each sample separately and normalize (using same method for all samples)

## Step 8:logNormalize

B73_ear_tip.list[[1]]<- NormalizeData(B73_ear_tip.list[[1]], normalization.method = "LogNormalize", scale.factor = 10000)

B73_ear_tip.list[[2]]<- NormalizeData(B73_ear_tip.list[[2]], normalization.method = "LogNormalize", scale.factor = 10000)

B73_ear_tip.list[[3]]<- NormalizeData(B73_ear_tip.list[[3]], normalization.method = "LogNormalize", scale.factor = 10000)

B73_ear_tip.list[[4]]<- NormalizeData(B73_ear_tip.list[[4]], normalization.method = "LogNormalize", scale.factor = 10000)


##Step 9:Find variables


B73_ear_tip.list[[1]]<- FindVariableFeatures(B73_ear_tip.list[[1]], selection.method = "vst", nfeatures = 2000)

B73_ear_tip.list[[2]]<- FindVariableFeatures(B73_ear_tip.list[[2]], selection.method = "vst", nfeatures = 2000)

B73_ear_tip.list[[3]]<- FindVariableFeatures(B73_ear_tip.list[[3]], selection.method = "vst", nfeatures = 2000)

B73_ear_tip.list[[4]]<- FindVariableFeatures(B73_ear_tip.list[[4]], selection.method = "vst", nfeatures = 2000)


# select features that are repeatedly variable across datasets for integration


features <- SelectIntegrationFeatures(object.list = B73_ear_tip.list)

```



## Next Perform integration
## We then identify anchors using the FindIntegrationAnchors() function, which takes a list of Seurat objects as input, and use these anchors to integrate all input datasets together with IntegrateData().

```{r}

# find anchors

anchors <- FindIntegrationAnchors(object.list = B73_ear_tip.list, anchor.features = features)


# Integrate data
# this command creates an 'integrated' data assay

B73_ear_tip.int <- IntegrateData(anchorset = anchors)

```


## Perform an integrated analysis
## Now we can run a single integrated analysis on all cells!

```{r}
# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay

DefaultAssay(B73_ear_tip.int)<-"integrated"


# Run the standard workflow for visualization and clustering
B73_ear_tip.int <- ScaleData(B73_ear_tip.int, verbose = FALSE)
B73_ear_tip.int <- RunPCA(B73_ear_tip.int, npcs = 30, verbose = FALSE)


```



```{r}
## Optional step: Determine the ‘dimensionality’ of the dataset
## Can also use default ndims = 30. Because the results did not change dramatically. 

ElbowPlot(B73_ear_tip.int, ndims = 40)

## Elbow plot: quantitative approach
## https://hbctraining.github.io/scRNA-seq/lessons/elbow_plot_metric.html

## Based on ElbowPlot(object, ndims = 40) plot, we could roughly determine the majority of the variation by where the elbow occurs (touches the ground). While this gives us a good rough idea of the number of PCs needed to be included, a more quantitative approach may be a bit more reliable. We can calculate where the principal components start to elbow by taking the larger (should be smaller?) value of:

## 1. The point where the principal components only contribute 5% of standard deviation and the principal components cumulatively contribute 90% of the standard deviation.
## 2. The point where the percent change in variation between the consecutive PCs is less than 0.1%.

## We will start by calculating the first metric:
# Determine percent of variation associated with each PC
pct <- B73_ear_tip.int[["pca"]]@stdev / sum(B73_ear_tip.int[["pca"]]@stdev) * 100

# Calculate cumulative percents for each PC
cumu <- cumsum(pct)

# Determine which PC exhibits cumulative percent greater than 90% and % variation associated with the PC as less than 5
co1 <- which(cumu > 90 & pct < 5)[1]

co1

## [1] result here (i.e. 43)

## The first metric returns PC (i.e. 43) as the PC matching these requirements. Let’s check the second metric, which identifies the PC where the percent change in variation between consecutive PCs is less than 0.1%:

# Determine the difference between variation of PC and subsequent PC
co2 <- sort(which((pct[1:length(pct) - 1] - pct[2:length(pct)]) > 0.1), decreasing = T)[1] + 1

# last point where change of % of variation is more than 0.1%.
co2

## [1] result here (i.e. 15)

## This second metric returns PC (i.e. 15). Usually, we would choose the minimum of these two metrics as the PCs covering the majority of the variation in the data.

# Minimum of the two calculation
pcs <- min(co1, co2)

pcs


```



```{r}
B73_ear_tip.int <- RunUMAP(B73_ear_tip.int, reduction = "pca", dims = 1:19) ## Use the PC number based on above Elbow plot: quantitative approach

B73_ear_tip.int <- FindNeighbors(B73_ear_tip.int, reduction = "pca", dims = 1:19)

B73_ear_tip.int <- FindClusters(B73_ear_tip.int, resolution = 0.85)

# Visualization

DimPlot(B73_ear_tip.int, reduction = "umap", label = TRUE, pt.size = 1)
DimPlot(B73_ear_tip.int, reduction = "umap", group.by = "orig.ident")

## two plots next to each other 

p1 <- DimPlot(B73_ear_tip.int, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(B73_ear_tip.int, reduction = "umap", label = TRUE, repel = TRUE)
p1 + p2

### split the UMAPs by samples 

DimPlot(B73_ear_tip.int, reduction = "umap", split.by = "orig.ident", pt.size = 1)


### group the UMAPs by samples. 

DimPlot(B73_ear_tip.int, reduction = "umap", group.by = "orig.ident", pt.size = 1)

```


```{r}
## plot UMIs and Genes on the UMAP

feature.pal = rev(colorRampPalette(brewer.pal(11,"Spectral"))(50))

FeaturePlot(object = B73_ear_tip.int, features = c("nCount_RNA"), cols = feature.pal, pt.size=1, reduction = "umap") + theme(axis.title.x=element_blank(),axis.title.y=element_blank(),axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank())

FeaturePlot(object = B73_ear_tip.int, features = c("nFeature_RNA"), cols = feature.pal, pt.size=1, reduction = "umap") + theme(axis.title.x=element_blank(),axis.title.y=element_blank(),axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank())
```


```{r}
# plot number of gene per cell on this plot. 
# https://github.com/satijalab/seurat/issues/2192
VlnPlot(B73_ear_tip.int, features = c("nFeature_RNA"), pt.size = 0.5)
```


```{r}
# plot number of transcript/UMI per cell on this plot. 
# https://github.com/satijalab/seurat/issues/2192
VlnPlot(B73_ear_tip.int, features = c("nCount_RNA"), pt.size = 0.5)
```


```{r}

VlnPlot(B73_ear_tip.int, features = c("nFeature_RNA", "nCount_RNA"), ncol=2)

```


```{r}
# Step 18: with cluster# and cell#
cell.num <- table(B73_ear_tip.int@active.ident)
ClusterLabels = paste("Cluster", names(cell.num), paste0("(n = ", cell.num, ")"))
ClusterBreaks = names(cell.num)

new.cluster.ids <- c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11","12","13","14","15")
names(new.cluster.ids) <- levels(B73_ear_tip.int)
scrna.int<-RenameIdents(B73_ear_tip.int, new.cluster.ids)
DimPlot(B73_ear_tip.int,reduction = "umap", label = TRUE, pt.size = 1, label.size = 10) + scale_color_discrete(breaks= ClusterBreaks, labels=ClusterLabels)


```



```{r}
## Step 20:save 

saveRDS(B73_ear_tip.int, file = "B73_ear_tip.int.rds")

### in future, when open this datase

B73_ear_tip.int <- readRDS(file = "B73_ear_tip.int.rds")


```


```{r}
## Remove cluster 3 with very low UMI (dead cells)

B73_ear_tip.int_remove_dead <- subset(B73_ear_tip.int, idents = 2, invert = TRUE)

DimPlot(B73_ear_tip.int_remove_dead, reduction = "umap", label = TRUE, pt.size = 1)


### split the integrated object to re-normalize and process

B73_ear_tip.list <- SplitObject(B73_ear_tip.int_remove_dead, split.by = "orig.ident")

B73_ear_tip.list



```{r}
## Before seurat intergration, filter each sample separately and normalize (using same method for all samples)


## need to set RNA assay before normalization and processing 

DefaultAssay(B73_ear_tip.list[[1]]) <- "RNA"
DefaultAssay(B73_ear_tip.list[[2]]) <- "RNA"
DefaultAssay(B73_ear_tip.list[[3]]) <- "RNA"
DefaultAssay(B73_ear_tip.list[[4]]) <- "RNA"



## Step 8:logNormalize

B73_ear_tip.list[[1]]<- NormalizeData(B73_ear_tip.list[[1]], normalization.method = "LogNormalize", scale.factor = 10000)

B73_ear_tip.list[[2]]<- NormalizeData(B73_ear_tip.list[[2]], normalization.method = "LogNormalize", scale.factor = 10000)

B73_ear_tip.list[[3]]<- NormalizeData(B73_ear_tip.list[[3]], normalization.method = "LogNormalize", scale.factor = 10000)

B73_ear_tip.list[[4]]<- NormalizeData(B73_ear_tip.list[[4]], normalization.method = "LogNormalize", scale.factor = 10000)


##Step 9:Find variables


B73_ear_tip.list[[1]]<- FindVariableFeatures(B73_ear_tip.list[[1]], selection.method = "vst", nfeatures = 2000)

B73_ear_tip.list[[2]]<- FindVariableFeatures(B73_ear_tip.list[[2]], selection.method = "vst", nfeatures = 2000)

B73_ear_tip.list[[3]]<- FindVariableFeatures(B73_ear_tip.list[[3]], selection.method = "vst", nfeatures = 2000)

B73_ear_tip.list[[4]]<- FindVariableFeatures(B73_ear_tip.list[[4]], selection.method = "vst", nfeatures = 2000)


# select features that are repeatedly variable across datasets for integration


features <- SelectIntegrationFeatures(object.list = B73_ear_tip.list)


```


## Next Perform integration
## We then identify anchors using the FindIntegrationAnchors() function, which takes a list of Seurat objects as input, and use these anchors to integrate all input datasets together with IntegrateData().

```{r}

# find anchors

anchors <- FindIntegrationAnchors(object.list = B73_ear_tip.list, anchor.features = features)


# Integrate data
# this command creates an 'integrated' data assay

B73_ear_tip.int <- IntegrateData(anchorset = anchors)

```


## Perform an integrated analysis
## Now we can run a single integrated analysis on all cells!

```{r}
# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay

DefaultAssay(B73_ear_tip.int)<-"integrated"


# Run the standard workflow for visualization and clustering
B73_ear_tip.int <- ScaleData(B73_ear_tip.int, verbose = FALSE)
B73_ear_tip.int <- RunPCA(B73_ear_tip.int, npcs = 30, verbose = FALSE)


```



```{r}
## Optional step: Determine the ‘dimensionality’ of the dataset
## Can also use default ndims = 30. Because the results did not change dramatically. 

ElbowPlot(B73_ear_tip.int, ndims = 40)

## Elbow plot: quantitative approach
## https://hbctraining.github.io/scRNA-seq/lessons/elbow_plot_metric.html

## Based on ElbowPlot(object, ndims = 40) plot, we could roughly determine the majority of the variation by where the elbow occurs (touches the ground). While this gives us a good rough idea of the number of PCs needed to be included, a more quantitative approach may be a bit more reliable. We can calculate where the principal components start to elbow by taking the larger (should be smaller?) value of:

## 1. The point where the principal components only contribute 5% of standard deviation and the principal components cumulatively contribute 90% of the standard deviation.
## 2. The point where the percent change in variation between the consecutive PCs is less than 0.1%.

## We will start by calculating the first metric:
# Determine percent of variation associated with each PC
pct <- B73_ear_tip.int[["pca"]]@stdev / sum(B73_ear_tip.int[["pca"]]@stdev) * 100

# Calculate cumulative percents for each PC
cumu <- cumsum(pct)

# Determine which PC exhibits cumulative percent greater than 90% and % variation associated with the PC as less than 5
co1 <- which(cumu > 90 & pct < 5)[1]

co1

## [1] result here (i.e. 43)

## The first metric returns PC (i.e. 43) as the PC matching these requirements. Let’s check the second metric, which identifies the PC where the percent change in variation between consecutive PCs is less than 0.1%:

# Determine the difference between variation of PC and subsequent PC
co2 <- sort(which((pct[1:length(pct) - 1] - pct[2:length(pct)]) > 0.1), decreasing = T)[1] + 1

# last point where change of % of variation is more than 0.1%.
co2

## [1] result here (i.e. 15)

## This second metric returns PC (i.e. 15). Usually, we would choose the minimum of these two metrics as the PCs covering the majority of the variation in the data.

# Minimum of the two calculation
pcs <- min(co1, co2)

pcs


```



```{r}
B73_ear_tip.int <- RunUMAP(B73_ear_tip.int, reduction = "pca", dims = 1:24) ## Use the PC number based on above Elbow plot: quantitative approach

B73_ear_tip.int <- FindNeighbors(B73_ear_tip.int, reduction = "pca", dims = 1:24)

B73_ear_tip.int <- FindClusters(B73_ear_tip.int, resolution = 0.65)

# Visualization

DimPlot(B73_ear_tip.int, reduction = "umap", label = TRUE, pt.size = 1)

DimPlot(B73_ear_tip.int, reduction = "umap", group.by = "orig.ident")

## two plots next to each other 

p1 <- DimPlot(B73_ear_tip.int, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(B73_ear_tip.int, reduction = "umap", label = TRUE, repel = TRUE)
p1 + p2

### split the UMAPs by samples 

DimPlot(B73_ear_tip.int, reduction = "umap", split.by = "orig.ident", pt.size = 1)


### group the UMAPs by samples. 

DimPlot(B73_ear_tip.int, reduction = "umap", group.by = "orig.ident", pt.size = 1)

```


```{r}
## plot UMIs and Genes on the UMAP

feature.pal = rev(colorRampPalette(brewer.pal(11,"Spectral"))(50))

FeaturePlot(object = B73_ear_tip.int, features = c("nCount_RNA"), cols = feature.pal, pt.size=1, reduction = "umap") + theme(axis.title.x=element_blank(),axis.title.y=element_blank(),axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank())

FeaturePlot(object = B73_ear_tip.int, features = c("nFeature_RNA"), cols = feature.pal, pt.size=1, reduction = "umap") + theme(axis.title.x=element_blank(),axis.title.y=element_blank(),axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank())
```


```{r}
# plot number of gene per cell on this plot. 
# https://github.com/satijalab/seurat/issues/2192
VlnPlot(B73_ear_tip.int, features = c("nFeature_RNA"), pt.size = 0.5)
```


```{r}
# plot number of transcript/UMI per cell on this plot. 
# https://github.com/satijalab/seurat/issues/2192
VlnPlot(B73_ear_tip.int, features = c("nCount_RNA"), pt.size = 0.5)
```


```{r}

VlnPlot(B73_ear_tip.int, features = c("nFeature_RNA", "nCount_RNA"), ncol=2)

```


```{r}
# UMAP cluster# and cell#
cell.num <- table(B73_ear_tip.int@active.ident)
ClusterLabels = paste("Cluster", names(cell.num), paste0("(n = ", cell.num, ")"))
ClusterBreaks = names(cell.num)

new.cluster.ids <- c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11","12","13","14","15")
names(new.cluster.ids) <- levels(B73_ear_tip.int)
scrna.int<-RenameIdents(B73_ear_tip.int, new.cluster.ids)
DimPlot(B73_ear_tip.int,reduction = "umap", label = TRUE, pt.size = 1, label.size = 10) + scale_color_discrete(breaks= ClusterBreaks, labels=ClusterLabels)


```



```{r}
## Step 20:save 

saveRDS(B73_ear_tip.int, file = "B73_ear_tip_int.rmDead.recluster.rds")

### in future, when open this datase

B73_ear_tip.int <- readRDS(file = "B73_ear_tip_int.rmDead.recluster.rds")


```



```{r}
## Prepare Subcluster6 by assigning graph.name="mygraph"

##next run UMAP

DefaultAssay(B73_ear_tip.int)<-"integrated"

B73_ear_tip.int.subcluster <- FindNeighbors(B73_ear_tip.int, graph.name= "mygraph", dims = 1:18)

B73_ear_tip.int.subcluster <- FindClusters(B73_ear_tip.int.subcluster, graph.name = "mygraph", resolution = 0.8)

B73_ear_tip.int.subcluster <- RunUMAP(B73_ear_tip.int.subcluster, reduction = "pca", dims = 1:30)


DimPlot(B73_ear_tip.int.subcluster, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 1, label.size = 10)


## save this object
saveRDS(B73_ear_tip.int.subcluster, file = "B73_ear_tip.int.subcluster.rds")


```


```{r}
```{r}
### Merge cell cycle G2/M clusters 

new.cluster.ids <- c("0", "1", "2", "3", "4", "5", "6", "7", "8", "7", "10", "11","12","7","14")
names(new.cluster.ids) <- levels(B73_ear_tip.int.subcluster)
B73_ear_tip.int.subcluster_mergeG2M <- RenameIdents(B73_ear_tip.int.subcluster, new.cluster.ids)
DimPlot(B73_ear_tip.int.subcluster_mergeG2M, reduction = "umap", label = TRUE, label.size = 4, pt.size = 1)

saveRDS(B73_ear_tip.int.subcluster_mergeG2M, file = "B73_ear_tip.int.subcluster_mygraph_mergeG2M.rds")


```



```{r}
### subcluster cluster
B73_ear_tip.int.subcluster6_mergeG2M <- FindSubCluster(B73_ear_tip.int.subcluster_mergeG2M, "6", "mygraph", subcluster.name = "subcluster6",  resolution = 0.2, algorithm = 1)


## Plot the sub-cluster 

DimPlot(B73_ear_tip.int.subcluster6_mergeG2M, reduction = "umap", group.by = "subcluster6", label = TRUE, label.size = 10, pt.size = 1)

DefaultAssay(B73_ear_tip.int.subcluster6_mergeG2M) <- "RNA"
VlnPlot(B73_ear_tip.int.subcluster6_mergeG2M, features = c("GRMZM2G529859"), group.by = "subcluster6") ##YAB8
VlnPlot(B73_ear_tip.int.subcluster6_mergeG2M, features = c("GRMZM2G014729"), group.by = "subcluster6") ##RA3-ok
VlnPlot(B73_ear_tip.int.subcluster6_mergeG2M, features = c("GRMZM2G003927"), group.by = "subcluster6") ##RA1
VlnPlot(B73_ear_tip.int.subcluster6_mergeG2M, features = c("GRMZM2G151044"), group.by = "subcluster6") ##ZmTPP4



## change cluster colors == what we used in Dev cell paper color 

DimPlot(B73_ear_tip.int.subcluster6_mergeG2M, reduction = "umap", group.by = "subcluster6", label = TRUE, label.size = 8, pt.size = 1,cols = c("1"="#afbca5","4"="#99b6ca","0"="#95511f","5"="#e761d8","10"="#98fb98","2"="#66dcf6","3"="#e93f33","6_1"="#64ac4e","6_0"="#f3f550","7"="#737172","8"="#e9967a","11"="#808000","12"="#4d6dcb","14"="#ffa500"))


## Cluster UMAP no label

DimPlot(B73_ear_tip.int.subcluster6_mergeG2M, reduction = "umap", group.by = "subcluster6", pt.size = 1,cols = c("1"="#afbca5","4"="#99b6ca","0"="#95511f","5"="#e761d8","10"="#98fb98","2"="#66dcf6","3"="#e93f33","6_1"="#64ac4e","6_0"="#f3f550","7"="#737172","8"="#e9967a","11"="#808000","12"="#4d6dcb","14"="#ffa500"))


DimPlot(B73_ear_tip.int.subcluster6_mergeG2M, reduction = "umap", group.by = "orig.ident", pt.size = 1)


DimPlot(B73_ear_tip.int.subcluster6_mergeG2M, reduction = "umap", split.by = "orig.ident", pt.size = 1)



saveRDS(B73_ear_tip.int.subcluster6_mergeG2M, file = "B73_ear_tip.int.subcluster6_mergeG2M.final.rds")


```


```{r}
## step 21: generate UMAPs for known markers. 


DefaultAssay(B73_ear_tip.int.subcluster6_mergeG2M)<-"RNA"


VlnPlot(B73_ear_tip.int.subcluster6_mergeG2M, features = c("GRMZM2G017087", "GRMZM2G372364", "GRMZM2G434380", "GRMZM2G165836", "GRMZM2G047448", "GRMZM2G028622", "GRMZM2G133331", "GRMZM2G059392", "GRMZM2G130953", "GRMZM2G102161", "GRMZM2G001289", "GRMZM2G101958", "GRMZM2G073671","GRMZM2G310115","GRMZM2G305046","GRMZM2G071959","GRMZM2G043854","GRMZM2G072342","GRMZM2G052544","GRMZM2G132794","GRMZM2G102218","GRMZM2G074543","GRMZM2G005353","GRMZM2G003927","GRMZM2G014729","GRMZM2G151044","GRMZM2G397518","GRMZM2G110242","GRMZM2G072274","GRMZM2G307119","GRMZM5G862109","GRMZM2G158717","GRMZM2G111696","GRMZM2G156016","GRMZM2G449165"))




FeaturePlot(B73_ear_tip.int.subcluster6_mergeG2M, features = c("GRMZM2G017087", "GRMZM2G001289", "GRMZM2G083725", "GRMZM2G372364", "GRMZM2G165836", "GRMZM2G047448", "GRMZM2G466532", "GRMZM2G354253", "GRMZM2G073671", "GRMZM2G310115", "GRMZM2G005619", "GRMZM2G305046", "GRMZM2G071959","GRMZM2G119071","GRMZM2G102161","GRMZM2G159397","GRMZM2G307119","GRMZM2G097989","GRMZM2G014729","GRMZM2G003927","GRMZM2G397518","GRMZM2G072274","GRMZM2G529859","GRMZM2G005353","GRMZM2G088309","GRMZM2G102218","GRMZM2G159399","GRMZM2G043854","GRMZM2G072342","GRMZM2G526668","GRMZM2G176141","GRMZM2G052544","GRMZM2G140694","GRMZM2G589696","GRMZM2G017470","GRMZM2G144172","GRMZM2G132794","GRMZM2G345700","GRMZM2G361652","GRMZM2G430849","GRMZM2G039074","GRMZM2G130953","GRMZM5G864847"))

"GRMZM2G017087", KN-1",
"GRMZM2G001289", HDZIV6
"GRMZM2G083725", LTP2
"GRMZM2G372364", CLE7
"GRMZM2G165836", FCP1
"GRMZM2G047448", WUS1
"GRMZM2G466532", Zmcle2_1
"GRMZM2G354253", Zmcle2_2
"GRMZM2G073671", Zmcyc3
"GRMZM2G310115", Zmcyc7_4b
"GRMZM2G005619", Zmcycb1_2
"GRMZM2G305046", Zmhis2a1
"GRMZM2G071959", Zmhis2b1
"GRMZM2G119071", Zmhis2b2
"GRMZM2G102161", Zmmads8
"GRMZM2G159397", Zmmads6
"GRMZM2G307119", bd1
"GRMZM2G097989", ZmGST41
"GRMZM2G014729", RA3
"GRMZM2G003927", RA1
"GRMZM2G397518", BA1
"GRMZM2G072274", BAF1
"GRMZM2G529859", Zmyab8
"GRMZM2G005353", Zmyab14
"GRMZM2G088309", drl1_Zmyab2
"GRMZM2G102218", drl2_Zmyab7
"GRMZM2G159399", ZmARF3_4_like
"GRMZM2G043854", Zmtmo5
"GRMZM2G072342", Zmtmo5_like1
"GRMZM2G526668", Zmtmo5_like2
"GRMZM2G176141", Zmtmo5_like3
"GRMZM2G052544", ZmAPL
"GRMZM2G140694", ZmDOF5.6
"GRMZM2G589696", ZmDOF5.6-LIKE
"GRMZM2G017470", ZmPEAR1_2_LIKE1
"GRMZM2G144172", ZmPEAR1_2_LIKE2
"GRMZM2G132794", ZmSHR1
"GRMZM2G345700", (2G345700), cortex
"GRMZM2G361652", ZmNPF6.4-LIKE2
"GRMZM2G430849", ZmNAC122, pith
"GRMZM2G039074", 2G039074 pith



```

```{r}
## export plots for BIF1 and BIF4 to show SPM and IM Central zone clusters. 

DotPlot(B73_ear_tip.int.subcluster6_mergeG2M, features = c("GRMZM2G130953","GRMZM5G864847"), idents = c("1","7"), dot.scale = 50)




```




```{r}
# total number of cells

B73_ear_tip.int.subcluster6_mergeG2M

An object of class Seurat 
30492 features across 11554 samples within 2 assays 
Active assay: RNA (28492 features, 0 variable features)
 1 other assay present: integrated
 2 dimensional reductions calculated: pca, umap
> 

# You can retrieve the number of cells expressing a gene my_gene from a Seurat object my_object in this way:

## https://github.com/satijalab/seurat/issues/371

##sum(GetAssayData(object = my_object, slot = "data")[my_gene,]>0)

##ZmWUS1 
sum(GetAssayData(object = B73_ear_tip.int.subcluster6_mergeG2M, slot = "data")["GRMZM2G047448",]>0)


##ZmCLE7
sum(GetAssayData(object = B73_ear_tip.int.subcluster6_mergeG2M, slot = "data")["GRMZM2G372364",]>0)


```








```{r}
# Step 24 for exporting table of coexpression value(pearson correlation value) for gene of interest, 

matrix <- B73_ear_tip.int.subcluster6_mergeG2M$RNA@data
matrix_mod<-as.matrix(matrix)

## for any gene of interest. pearson correlation 
gene<-as.numeric(matrix_mod["GRMZM2G372364",])
correlations<-apply(matrix_mod,1,function(x){cor(gene,x)})
write.table(correlations, file = "B73_ear_tip.int.subcluster6_mergeG2M_CLE7_Pearson_correlation.txt")


## for any gene of interest. spearman correlation 
gene<-as.numeric(matrix_mod["GRMZM2G372364",])
correlations<-apply(matrix_mod,1,function(x){cor(gene,x, method = "spearman")})
write.table(correlations, file = "B73_ear_tip.int.subcluster6_mergeG2M_CLE7_Spearman_correlation.txt")


# for generating the distribution and plot any gene's position on the distribution

B73_tip_cle7=read.table("B73_ear_tip.int.subcluster6_mergeG2M_CLE7_Spearman_correlation.txt")

f=B73_tip_cle7[,1]!=1

hist(B73_tip_cle7[f,1],breaks = 1000,main = "B73_tip_Spearman_correlation",xlab="Coexpression")

# to plot position of WUS1 on CLE7 coexpression's distribution

abline(v=B73_tip_cle7["GRMZM2G047448",1],col="red",lwd=2,lty=2)

# to plot position of LOG7 on CLE7 coexpression's distribution
abline(v=B73_tip_cle7["GRMZM2G059392",1],col="blue",lwd=2,lty=2)

# to plot position of FEA4 on CLE7 coexpression's distribution
abline(v=B73_tip_cle7["GRMZM2G133331",1],col="orange",lwd=2,lty=2)



# for generating the distribution and plot any gene's position on the distribution

B73_tip_cle7=read.table("B73_ear_tip.int.subcluster6_mergeG2M_CLE7_Pearson_correlation.txt")

f=B73_tip_cle7[,1]!=1

hist(B73_tip_cle7[f,1],breaks = 1000,main = "B73_tip_Pearson_correlation",xlab="Coexpression")

# to plot position of WUS1 on CLE7 coexpression's distribution

abline(v=B73_tip_cle7["GRMZM2G047448",1],col="red",lwd=2,lty=2)

# to plot position of LOG7 on CLE7 coexpression's distribution
abline(v=B73_tip_cle7["GRMZM2G059392",1],col="blue",lwd=2,lty=2)

# to plot position of FEA4 on CLE7 coexpression's distribution
abline(v=B73_tip_cle7["GRMZM2G133331",1],col="orange",lwd=2,lty=2)



```



```{r}
## FindAllMarkers

all.markers <- FindAllMarkers(object = B73_ear_tip.int.subcluster6_mergeG2M)

write.table(all.markers, file=sprintf("B73_ear_tip.int.subcluster6_mergeG2M_FindAllMarkers.xls"), quote=FALSE, sep="\t", row.names=FALSE);

```


```{r}
# Step 25: export the top10 markers heatmap in each cluster, use lower cutoff https://github.com/satijalab/seurat/issues/1510

## first need to fun default assay as integrated, because when generate UMAP, we used RNA assay. 

DefaultAssay(B73_ear_tip.int.subcluster6_mergeG2M)<-"integrated"

object5.markers <- FindAllMarkers(object = B73_ear_tip.int.subcluster6_mergeG2M, features = VariableFeatures(B73_ear_tip.int.subcluster6_mergeG2M, assay = 'integrated'),
                                                                                only.pos = TRUE, 
                                                                                min.pct = 0.25, 
                                                                                logfc.threshold = 0.25, 
                                                                                assay = 'integrated', 
                                                                                slot = c('data'))

write.table(object5.markers, file=sprintf("B73_ear_tip.int.subcluster6_mergeG2M_minpct0.25.Wilcox.xls"), quote=FALSE, sep="\t", row.names=FALSE);


object5.markers %>% group_by(cluster) %>% top_n(n = 2, wt = avg_logFC)


## generate top50 genes
top50 <- object5.markers %>% group_by(cluster) %>% top_n(n = 50, wt = avg_logFC)

DoHeatmap(object = B73_ear_tip.int.subcluster6_mergeG2M, 
          features = top50$gene, 
          disp.min = -1, 
          disp.max = 1) + scale_fill_gradientn(colors = colorRampPalette(c("#0200ad", 
                                                                           "#fbfcbd", 
                                                                           "#ff0000"))(256))
## generate top10 genes

top10 <- object5.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)

DoHeatmap(object = B73_ear_tip.int.subcluster6_mergeG2M, 
          features = top10$gene, 
          disp.min = -1, 
          disp.max = 1) + scale_fill_gradientn(colors = colorRampPalette(c("#0200ad", 
                                                                           "#fbfcbd", 
                                                                           "#ff0000"))(256))


```


```{r}

## subcluster IM and SPM to run slingshot trajectory. 


## IM use the original scale 

IM <- subset(B73.tip, idents = c(1, 3, 4))

IM.oriscale <- RunUMAP(IM, reduction = "pca", dims = 1:30) ## Use default 30, or change the PC number based on above Elbow plot: quantitative approach

# Visualization

DimPlot(IM.oriscale, reduction = "umap", label = TRUE, pt.size = 1)

DimPlot(IM.oriscale, reduction = "umap", group.by = "subcluster6", pt.size = 1, label = TRUE, )


## Save RDS

saveRDS(IM.oriscale, file = "IM.oriscale.rds")

setwd("C:/Users/Xiaosa Xu/Desktop/SC Stem Cell Manuscript/Seurat_sc_datasets/1_B73_ear_tip_xu16192932_Dec2022_integration/Trajectory")

IM.oriscale <- readRDS(file="IM.oriscale.rds")


# Slingshot
# https://github.com/kstreet13/slingshot/issues/105

sdsIM.ori <- slingshot(Embeddings(IM.oriscale, "umap"), clusterLabels = IM.oriscale$subcluster6,
start.clus = 3, end.clus = c("1","4"), stretch = 0)

sdsIM.ori

## Unfortunately, slingshot does not natively support ggplot2. So this is a function that assigns colors to each cell in base R graphics.
## https://bustools.github.io/BUS_notebooks_R/slingshot.html#slingshot
## https://bustools.github.io/BUS_notebooks_R/slingshot.html#trajectory_inference


#' Assign a color to each cell based on some value
#' 
#' @param cell_vars Vector indicating the value of a variable associated with cells.
#' @param pal_fun Palette function that returns a vector of hex colors, whose
#' argument is the length of such a vector.
#' @param ... Extra arguments for pal_fun.
#' @return A vector of hex colors with one entry for each cell.

cell_pal <- function(cell_vars, pal_fun,...) {
  if (is.numeric(cell_vars)) {
    pal <- pal_fun(100, ...)
    return(pal[cut(cell_vars, breaks = 100)])
  } else {
    categories <- sort(unique(cell_vars))
    pal <- setNames(pal_fun(length(categories), ...), categories)
    return(pal[cell_vars])
  }
}



cell_colors_clust <- cell_pal(IM.oriscale$subcluster6, hue_pal()) 

plot(reducedDim(sdsIM.ori), col = cell_colors_clust, pch = 16, cex = 1)

lines(sdsIM.ori, lwd = 2, type = 'c', col = 'black') ## make sure type = 'c' which is the curve. 
lines(sdsIM.ori, lwd = 2, type = 'lineage', col = 'black') ## make sure type = 'c' which is the curve.


## Plot pseudotime for gene of interest. 

exp<-GetAssayData(IM.ori, slot = "data")

plotGenePseudotime(sdsIM.ori, "GRMZM2G372364", exp ) ## plot gene of interest one at a time for each lineages

plotGenePseudotime(sdsIM.ori, "GRMZM2G372364", col = cell_colors_clust, exp ) ## plot gene of interest one at a time for each lineages


## Which cells are in which lineage? Here we plot the pseudotime values for each lineage.


### curve
nc <- 3

pt <- slingPseudotime(sdsIM.ori)

nms <- colnames(pt)

nr <- ceiling(length(nms)/nc)

pal <- viridis(100, end = 0.95)

par(mfrow = c(nr, nc))

for (i in nms) {
  colors <- pal[cut(pt[,i], breaks = 100)]
  plot(reducedDim(sdsIM.ori), col = colors, pch = 16, cex = 1, main = i)
  lines(sdsIM.ori, lwd = 2, col = 'black', type = 'c')
}



#### lineage
nc <- 3

pt <- slingPseudotime(sdsIM.ori)

nms <- colnames(pt)

nr <- ceiling(length(nms)/nc)

pal <- viridis(100, end = 0.95)

par(mfrow = c(nr, nc))

for (i in nms) {
  colors <- pal[cut(pt[,i], breaks = 100)]
  plot(reducedDim(sdsIM.ori), col = colors, pch = 16, cex = 1, main = i)
  lines(sdsIM.ori, lwd = 2, col = 'black', type = 'lineage')
}


## plot pseudotime across all clusters. 


```





```{r}

## subcluster IM and SPM to run slingshot trajectory. 


## SPM

SPM.ori.scale <- subset(B73.tip, idents = c(5, 6, 12, 14))


SPM.ori.scale <- RunUMAP(SPM.ori.scale, reduction = "pca", dims = 1:30) ## Use default 30, or change the PC number based on above Elbow plot: quantitative approach

# Visualization

DimPlot(SPM.ori.scale, reduction = "umap", label = TRUE, pt.size = 1)

DimPlot(SPM.ori.scale, reduction = "umap", group.by = "subcluster6", pt.size = 1, label = TRUE)


## Save RDS

saveRDS(SPM.ori.scale, file = "SPM.ori.scale.rds")

setwd("C:/Users/Xiaosa Xu/Desktop/SC Stem Cell Manuscript/Seurat_sc_datasets/1_B73_ear_tip_xu16192932_Dec2022_integration/Trajectory")

SPM.ori.scale <- readRDS(file="SPM.ori.scale.rds")


# Slingshot
# https://github.com/kstreet13/slingshot/issues/105

sdsSPM.ori <- slingshot(Embeddings(SPM.ori.scale, "umap"), clusterLabels = SPM.ori.scale$subcluster6,
start.clus = 5, end.clus = c("6_0","6_1","12","14"), stretch = 0)

sdsSPM.ori

## Unfortunately, slingshot does not natively support ggplot2. So this is a function that assigns colors to each cell in base R graphics.
## https://bustools.github.io/BUS_notebooks_R/slingshot.html#slingshot
## https://bustools.github.io/BUS_notebooks_R/slingshot.html#trajectory_inference


#' Assign a color to each cell based on some value
#' 
#' @param cell_vars Vector indicating the value of a variable associated with cells.
#' @param pal_fun Palette function that returns a vector of hex colors, whose
#' argument is the length of such a vector.
#' @param ... Extra arguments for pal_fun.
#' @return A vector of hex colors with one entry for each cell.

cell_pal <- function(cell_vars, pal_fun,...) {
  if (is.numeric(cell_vars)) {
    pal <- pal_fun(100, ...)
    return(pal[cut(cell_vars, breaks = 100)])
  } else {
    categories <- sort(unique(cell_vars))
    pal <- setNames(pal_fun(length(categories), ...), categories)
    return(pal[cell_vars])
  }
}



cell_colors_clust <- cell_pal(SPM.ori.scale$subcluster6, hue_pal()) 

plot(reducedDim(sdsSPM.ori), col = cell_colors_clust, pch = 16, cex = 1)

lines(sdsSPM.ori, lwd = 2, type = 'c', col = 'black') ## make sure type = 'c' which is the curve. 
lines(sdsSPM.ori, lwd = 2, type = 'lineage', col = 'black') ## make sure type = 'c' which is the curve.

## Plot pseudotime for gene of interest. 

DefaultAssay(SPM.ori.scale) <- "RNA"

exp<-GetAssayData(SPM.ori.scale, slot = "data")

plotGenePseudotime(sdsSPM.ori, "GRMZM2G372364", exp ) ## plot gene of interest one at a time for each lineages

plotGenePseudotime(sdsSPM.ori, "GRMZM2G372364", col = cell_colors_clust, exp ) ## plot gene of interest one at a time for each lineages


## Which cells are in which lineage? Here we plot the pseudotime values for each lineage.


## curve
nc <- 3

pt <- slingPseudotime(sdsSPM.ori)

nms <- colnames(pt)

nr <- ceiling(length(nms)/nc)

pal <- viridis(100, end = 0.95)

par(mfrow = c(nr, nc))

for (i in nms) {
  colors <- pal[cut(pt[,i], breaks = 100)]
  plot(reducedDim(sdsSPM.ori), col = colors, pch = 16, cex = 1, main = i)
  lines(sdsSPM.ori, lwd = 2, col = 'black', type = 'c')
}


## lineage
nc <- 3

pt <- slingPseudotime(sdsSPM.ori)

nms <- colnames(pt)

nr <- ceiling(length(nms)/nc)

pal <- viridis(100, end = 0.95)

par(mfrow = c(nr, nc))

for (i in nms) {
  colors <- pal[cut(pt[,i], breaks = 100)]
  plot(reducedDim(sdsSPM.ori), col = colors, pch = 16, cex = 1, main = i)
  lines(sdsSPM.ori, lwd = 2, col = 'black', type = 'lineage')
}



## plot pseudotime across all clusters. 


```


